---
title: "Juvenile Snowshoe Hares - Bonanza Creek"
author: "Eli Schulman"
date: "11/3/2021"
output: html_document
---

***Introduction: ***

***Data and Analysis: *** 

***Data Citation: *** Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```


```{r}
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  clean_names()
```

### Annual Juvenile Hare Trap Coounts

```{r}
annual_traps <- bonanza_hares %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarize(hares_trapped = n())

ggplot(annual_traps, aes(x = year, y = hares_trapped)) +
  geom_bar(stat = "identity", 
           fill = "burlywood4",
           width = .5) +
  scale_x_discrete(limit = c(1998:2012)) +
  theme_minimal() +
  labs (title = "Hares Trapped at Bonanza Creek",
        y = "# of Hares",
        x = "Year",
        caption = "Juvenile Snowshoe Hares trapped at the Bonanza Creek Expirimental Forest") +
  theme(plot.title = element_text(hjust = .5)) 

trap_trends <- annual_traps %>%
  summarize(mean = round(mean(hares_trapped),0),
            max = max(hares_trapped),
            min = min(hares_trapped),
            standard_deviation = sd(hares_trapped))
```
Trapped Snowshoes Takeaway: There were, on average, 213 Juvenile Snowshoe Hares trapped each year during the study. However, there was substantial variability in the annual trappings (sd = 138). For example, trappings were lowest in 2002 (38 hares trapped) and peaked in 2009 (410 hares trapped). These swings are evidence of non-standard sampling efforts between years. Moving forward, the scientists should ensure that the traps used for the study remain consistent in frequency and time of year that they are monitored. 












